---
title: "turnover_time"
author: "Max Blasdel"
date: "April 26, 2019"
output: html_document
---
# Purpose
Process decay rate rasters to display turnover times

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load relavent packages
```{r}
library(raster)
```

Load data
```{r}
ras <- dir("../../outputs/rasters/", pattern = ".tif$", full.names = T)
ras_name <- dir("../../outputs/rasters/", pattern = ".tif$", full.names = F)
#ras_name <- tools::file_path_sans_ext(ras_name)
```


$$N_{t}=N_{o} * exp^{(-k*t)}$$
where

$$.5 = exp^{(-k*t)}$$
where

$$ln(.5)=-k*t$$

$$\frac{ln(.5)}{-k} = t$$
where k is know and is the values of the rasters

Function will create a temporary raster. Create a folder for this raster to go into to keep track of these temp files.
```{r}
for (i in 1:length(ras)) {
  raster <- raster(ras[i])
  #creates unique filepath for temp directory
  dir.create (file.path("temp"), showWarnings = FALSE)
  
  #sets temp directory
  rasterOptions(tmpdir=file.path("temp")) 
  
  time <- -log(.05)/raster # set for t95
  
  writeRaster(time, paste0("time/t95_", ras_name[i]))
  
  unlink("temp", recursive = T)
}
```

Reference simplistic approach to decay given a constant k value across sites

Come back to this and work on a bit more
```{r}
df <- data.frame(
  "t50_cwd" = -log(0.5)/0.05,
  "t95_cwd"  = -log(0.05)/0.05,
  "t50_fwd"  = -log(0.5)/0.09,
  "t95_fwd"  = -log(0.05)/0.09
)

# Check 
1*exp(-0.05 * seq(1,60,1))
```


# advanced plotting in R, mayby not worth the time
```{r}
library(rgdal)
library(rasterVis)
library(RColorBrewer)
```

```{r}
colr <- colorRampPalette(brewer.pal(11, 'RdYlBu'))

levelplot(cwd_time, 
          margin=FALSE,                       # suppress marginal graphics
          colorkey=list(
            space='bottom',                   # plot legend at bottom
            labels=list(at=0:100, font=4)      # legend ticks and labels 
          ),    
          par.settings=list(
            axis.line=list(col='transparent') # suppress axes and legend outline
          ),
          scales=list(draw=FALSE),            # suppress axis labels
          col.regions=colr,                   # colour ramp
          at=seq(0, 100, len=101))             # colour ramp breaks
  #layer(sp.polygons(oregon, lwd=3))           # add oregon SPDF with latticeExtra::layer
```

