---
author: "Max Blasdel"
date: "Fenruary 24, 2019"
title: "Deriving missing values for FWD and CWD"
---

# Purpose
Create derived values for CWD and FWD from mean genus values with recalulated climate agnostic values

```{r}
library(tidyverse)
```

load data
```{r}
dat <- read.csv("../data/mean.species.k.csv")
```

```{r}
plot(dat$CWD, dat$FWD)
```

Applying derived values equation
$$y=0.0393 + 1.2535x$$
There is an issue here as SESE CWD becomes negative with this approach
```{r}
dat %<>% 
  mutate(der.FWD = 0.0393 + CWD*1.2535,
         der.CWD = (FWD-0.0393)/1.2535)
```

```{r}
dat %<>% 
  mutate(CWD = coalesce(CWD, der.CWD),
         FWD = coalesce(FWD, der.FWD))
dat %>% filter(genus == "Tsuga")
```

write out for later use
```{r}
write.csv(dat, "../data/mean_species_k_wDerived.csv")
```











Produces really good r values since there is modeled data in the regression
```{r}
lin_model <- lm(FWD ~ CWD, data = dat)
summary(lin_model)

{plot(dat$CWD, dat$FWD)
abline(lin_model)}
```


# old

Essentially the correlation test is giving poor results when comparing the recalculated values. I might just use the same equation that is in the framework document. I do have to recreate this script since it disappeared somehow. 
```{r}
cor(dat$CWD, dat$FWD, method = c("pearson", "kendall", "spearman"))

cor.test(dat$CWD, dat$FWD, method = "kendall")
cor.test(dat$CWD, dat$FWD, method = "pearson")
cor.test(dat$CWD, dat$FWD, method = "spearman")

dat_no_que <- dat[-3,] 

lin_model <- lm(FWD ~ CWD, data = dat_no_que)
summary(lin_model)


{plot(dat_no_que$CWD, dat_no_que$FWD)
abline(lin_model)}
```

removing na values and making sure only the CWD/FWD values are plotted against each other
```{r}
d <- dat_no_que %>% 
  na.omit()

lin_model <- lm(FWD ~ CWD, data = d)
summary(lin_model)

{plot(d$CWD, d$FWD)
abline(lin_model)}

cor.test(d$CWD, d$FWD, method = "pearson")
```

