---
title: "Climate Mod Adair and Smith"
date: "January 29, 2019"
author: "Max Blasdel"
---
*update 3/8/2019*
# Implement alternative temperature equations and create masked version of the climate modifier.

Implement climate models and clip to California. 

Need to confirm projection of data before clipping and writing out

```{r message=FALSE}
library(raster)
library(dplyr)
library(rgdal)
library(magrittr)
library(sf)
library(rgeos)
```

Climate equations to use
Adair functions
```{r}
# From Adair
FW_Adair <- function(PPT ,PET) {

  return(1 / (1 + 30 * exp((-8.5 * (PPT / PET)))))

}

FT_Adair <- function(Temp) {

  return(exp(308.56 * ((1 / 56.02) - (1 / ((273.15 + Temp) - 227.13)))))

}
```

Q10 equations
```{r}
FT_Q10 <- function(Temp, Q10) {
  return(Q10^((Temp-10)/10))
}
```

Smith - DONT USE - AET is a different metric than PET
```{r}
# # From Smith 2011
# Climate_Smith <- function(PPT, Temp_July) {
#   (PPT/1000)^0.5 * 1*exp(-66500/8.31 * ((1/Temp_July) - (1/15)))
# }
# Smith_simple <- function(AET) {
#   AET/1620
# }
```

Cropping functions
```{r}
# crop and mask function
cropAndMask <- function(raster, shp) {
  r <- crop(raster, shp)
  r <- mask(r, shp)
  return(r)
}
```

Read in data
```{r}
pet <- readRDS("../../../GRIDMET/ncFiles/pet")
pr <- readRDS("../../../GRIDMET/ncFiles/pr")
temp <- readRDS("../../../GRIDMET/ncFiles/Tmean")
#summary.data.frame(temp)
# another temp data file exists but it reports data in celcius not kelvin
#temp <- readRDS("../../../GRIDMET/ncFiles/Tmean_df")
```

Read in California shapefile to clip

```{r}
# Cali <- readOGR("../../data/California", layer = "CA_State_TIGER2016")
# 
# Cali <- spTransform(Cali, CRSobj = '+proj=longlat +datum=WGS84 +no_defs')

# gBuffer(Cali)
# 
Cali <- read_sf("../../data/California")

Cali <- st_transform(Cali, crs = 4326)

# apply buffer to shapefile to reduce data loss

Cali <- st_buffer(Cali, dist = 1)

Cali <- as(Cali, "Spatial")

plot(Cali)
```

Simple Smith models

```{r}
# #isolate the coordinates
# lonlat <- pet[,1:2]
# lonlat$TenYrAverage <- Smith_simple(pet$TenYrAverage)
# 
# # Convert data to raster format
# smith_rast <- raster::rasterFromXYZ(lonlat, crs ="+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
# smith_rast <- projectRaster(from = smith_rast, crs = coordinate_ref)
# 
# # Crop and mask
# smith_rast <- cropAndMask(raster = smith_rast, shp = Cali)
# 
# plot(smith_rast)
```

Adair model moisture

```{r}
lonlat <- pet[,1:2]

lonlat$climate <- FW_Adair(PPT = pr$TenYrAverage, PET = pet$TenYrAverage)

adair_fw <- raster::rasterFromXYZ(lonlat, crs ="+proj=longlat +datum=WGS84 +no_defs")

#adair_fw <- projectRaster(from = adair_fw, crs = coordinate_ref)

adair_fw <- cropAndMask(raster = adair_fw, shp = Cali)

#plot(adair_fw)
```

Adair temperature

```{r}
lonlat <- pet[,1:2]
lonlat$climate <- FT_Adair(Temp = temp$TenYrAverage)
adair_ft <- raster::rasterFromXYZ(lonlat, crs ="+proj=longlat +datum=WGS84 +no_defs")
#adair_ft <- projectRaster(from = adair_ft, crs = coordinate_ref)
adair_ft <- cropAndMask(raster = adair_ft, shp = Cali)

#plot(adair_ft)
```

Multiply together

```{r}
adair_climate <- adair_ft * adair_fw
```

Filter map to show areas of neutral change
```{r}
# useValues sets values outside clamp range to NA
mostly_neutral <- clamp(adair_climate, lower = .9, upper = 1.1, useValues = F)

# inspect
plot(mostly_neutral)

# output plot
# finish in GIS software
writeRaster(mostly_neutral, "../../climate_mod/mostly_neut.tif")
```


# write out 
# V2 was a rerun done in late June with the buffered shapefile of CA. The equation was also tweaked, changing 273.15 to 273. This slightly alters the values of the final output. After a successful run, correct clipping, this is being rerun as V3 with the 273.15 value. 

```{r}
writeRaster(adair_climate, "../../climate_mod/adair_V3.tif", overwrite = T)
```


# Below should be the same as above
Note the Adiar moisture function will stay the same

```{r}
lonlat <- pet[,1:2]
lonlat$moist <- FW_Adair(PPT = pr$TenYrAverage, PET = pet$TenYrAverage)
lonlat$temp <- FT_Adair(Temp = temp$TenYrAverage)

lonlat %<>% 
  mutate(climate = temp*moist)

lonlat <- lonlat[,c(1,2,5)]
climate_raster <- rasterFromXYZ(lonlat, crs = '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0')

climate_raster <- cropAndMask(climate_raster, Cali)

climate_raster
```

Produces slightly different values, probably some rounding error from when the clips happen? Does not look to be significantly different. 

Assuming above is correct, right out for later use.

```{r}
#writeRaster(climate_raster, "climate_adair.tif", overwrite = T)
```



```{r}
# create folder for rasters to test Q10 values
dir.create("Sensitivity")

vals <- seq(1.4, 2, .2)
for (i in vals) {
# extract x/y coordinates from a climate layer
lonlat <- pet[,1:2]
# run equation
lonlat$climate <- FT_Q10(Q10= i, Temp = temp$TenYrAverage)
# transform to raster, project and crop to cali
temp_rast <- raster::rasterFromXYZ(lonlat, crs ="+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
temp_rast <- projectRaster(from = temp_rast, crs = coordinate_ref)
temp_rast <- cropAndMask(raster = temp_rast, shp = Cali)

cli <- temp_rast * adair_fw
writeRaster(cli, filename = paste0("Sensitivity/Q10", i, ".tif"))
}
```







