---
title: "Climate By Species"
author: "Max Blasdel"
date: "2/12/2019"
---
# Purpose
Implement new approach to climate modifier. 
I want to start by deratifying the UW raster based on species. 
I'll try exporting these as individual rasters as opposed to csv files.

Load packages
```{r message=FALSE}
library(raster)
library(dplyr)
library(tidyr)
library(magrittr)
library(data.table)
```


I want to start by loading a raster of just FCID values and joining species distribution data.

*I think what happened is that when writing out the FCID values only some get copied over and there was a signicant portion that were lost. This is why there was no ABAM in the generated raster.* 
```{r}
# run once
FCID <- raster("E:/Drive_sync/CBI_Unshared_Work/CARBCAT Development/Data/UW Modeled Biomass Resource/Treatment Tiffs/Remove100Percent.tif")

#writeRaster(r, "FCID_UW.tif")
```

Load species distribution and proportion data
```{r}
# FCID <- raster(".tif")
data <- readRDS("database") %>% 
  dplyr::select(FCID2018, Species, proportion) %>% 
  rename(ID = FCID2018)
head(data)
```

Spread data into longer database. Number of rows will now match the FCIDs
```{r}
data %<>% 
  spread(Species, proportion)
data %<>% dplyr::select(-V1)

data %>% head()
```

Join with FCID raster
```{r}
levels(FCID)[[1]] <- data
levels(FCID)
```

# For reclassification
```{r}
#vals <- getValues(FCID) %>% unique() %>% desc()
```

Thinking through why ratify leaves out so many entries.
Try to reclass FCID values to less digits
```{r}
#reclassify(FCID, matrix(vals, seq(1, 226203, 1)))?
```


*2/13/19*
JUGLA last raster produced
*2/14/19*
PLRA last raster produced
*2/15/19*
TOCA last raser produced
*2/18/19* 
All rasters run
```{r}
names <- colnames(data)[2:ncol(data)]

names
names<-names[68:70]

# for (i in vector) {
#   rast <- deratify(FCID, names[i])
#   writeRaster(rast, paste("E:/rasterSpecies/indiv_rasts_UW", names[i], ".tif", sep = ""))
# }

  
names <- as.list(names)
# Create temp folder for rasters that get created
dir.create(file.path("E:/tempdir"), showWarnings = F)
rasterOptions(tmpdir = file.path("E:/tempdir"))

lapply(names, function(x) {
  deratify(FCID, x) %>% 
    writeRaster(paste("E:/rasterSpecies/indiv_rasts_UW/", x, ".tif", sep = ""))
})

```

# Update for angiosperm/gymnosperm distinction used in foliage

Read in data. 
```{r}
FCID <- raster("C:/Users/Max Blasdel/Desktop/Treatment Tiffs/Remove100Percent.tif")

dat <- fread("full_database_decay.csv") %>% 
  dplyr::select(FCID2018, Classification, proportion) %>% 
  rename(ID = FCID2018)
dat %>% head()
```


```{r eval=FALSE}
dat %>%
  spread(Classification, proportion)
# error becuase there are multiple ang/gym in a given FCID
```

Sum the prpoportions and then spread
```{r}
dat %<>%
  group_by(ID, Classification) %>% 
  mutate(new_prop = sum(proportion)) %>% 
  ungroup() %>% 
  dplyr::select(-proportion) %>% 
  distinct()

dat %<>% 
  spread(Classification, new_prop) %>% 
  dplyr::select(-V1)
dat
```

Needs to be coerced into a data table before proceeding
```{r}
dat %>% class()
dat2 <- as.data.table(dat)
dat2 %>% class()
FCID

levels(FCID)[[1]] <- dat2

levels(FCID)
```

# outputs Angiosperm and Gymnosperm proportion rasters

```{r}
names <- colnames(dat[2:ncol(dat)])
names <- as.list(names)

# Create temp folder for rasters that get created
dir.create(file.path("C:/Users/Max Blasdel/Desktop/tempdir"), showWarnings = F)
rasterOptions(tmpdir = file.path("C:/Users/Max Blasdel/Desktop/tempdir"))

lapply(names, function(x) {
  deratify(FCID, x) %>% 
    writeRaster(paste(x, ".tif", sep = ""))
})

```

