---
title: "Alternate Climate Equations"
date: " January 29, 2019"
author: "Max Blasdel"
---

# Purpose
Prepare data for Adair 2008 model and Smith 2011 model

```{r message=FALSE}
library(ncdf4)
library(dplyr)
library(data.table)
```


Adair 2008
Note the temp equation is the from Lloyd and Taylor
Equations will be run in other script
```{r}
FW_Adair <- function(PPT ,PET) {
 return(1/(1 + 30 * exp((-8.5 * (PPT/PET)))))
}
FT_Adair <- function(Temp) {
return(exp(308.56 * ((1/56.02)-(1/((273.15 + Temp)-227.13)))))
}
```

Smith et al. 2011
```{r}
Climate_Smith <- function(PPT, Temp_July) {
  (PPT/1000)^0.5 * 1*exp(-66500/8.31 * ((1/Temp_July) - (1/15)))
}
smith_simple <- function(AET) {
  AET/1620
}
```

Read in data for Adair model from nc files

First pull in latlon grid
```{r}
nc_list <- dir(path = "../../../../../gridMET Data/pet", pattern = "*\\.nc$", full.names = T)
nc<-nc_open(nc_list[1], verbose = F)
lon <- ncvar_get(nc, "lon")
lat <- ncvar_get(nc, "lat")
lonlat<-expand.grid(lon, lat)
rm(nc,lon,lat)
gc()
```

Relevent functions
```{r}
get_nc_variable<-function(nc_file, dname, fun){
  nc <- nc_open(nc_file, verbose = F) # open the nc file
  ncArray<-ncvar_get(nc, dname) # get values for variable of interest 
  array_yearly<-apply(ncArray, c(1,2), fun) # apply fun for entire year, apply in matrix
  yearly.vec <- as.vector(array_yearly) # put as vector
  df <- data.frame(yearly.vec) # put as dataframe
}
```

Variables changed
```{r}
# PET
dname <- "potential_evapotranspiration"
variable_path <- "pet"

# PPT
dname <- "precipitation_amount" # Need variable name
variable_path <- "pr"
```

# Change above variables
```{r}
nc_list <- dir(path = paste("../../../../../gridMET Data/", variable_path, sep=""), pattern = "*\\.nc$", full.names = T)

pet_list <- list()
for (i in 1:length(nc_list)) {
  pet_list[[i]]<-get_nc_variable(nc_list[i], dname = dname, fun = sum)
}
  # bind together
pet_list <- do.call(cbind, pet_list)
#columns need unique names to average together
colnames(pet_list) <- seq(2007, 2017, 1) #change when more yeasrs downloaded
# Take average of all years
TenYrAve <-pet_list %>%
  transmute(TenYrAverage = rowMeans(.)) %>%
  cbind(lonlat, .) %>%
  as.data.table()
TenYrAve
```

```{r}
saveRDS(TenYrAve, paste("../../ncFiles/Processed Data/", variable_path, sep = ""))
```

Dealing with temperature. I think I want to look at how the min and max were averaged together. I averaged entire min and max for ten(eleven) years and then averaged the two. Want to average each year.
```{r}
dname <- "air_temperature"
variable_path <- "Tmin"
```

Run for Min values first
Then run for Max
```{r}
nc_list <- dir(path = paste("../../../../../gridMET Data/", variable_path, sep=""), pattern = "*\\.nc$", full.names = T)
var_list <- list()
for (i in 1:length(nc_list)) {
  var_list[[i]]<-get_nc_variable(nc_list[i], dname = dname, fun = mean)
}
min<-do.call(cbind,var_list)

# Change path name, variable stays the same
variable_path <- "Tmax"

nc_list <- dir(path = paste("../../../../../gridMET Data/", variable_path, sep=""), pattern = "*\\.nc$", full.names = T)
var_list <- list()
for (i in 1:length(nc_list)) {
  var_list[[i]]<-get_nc_variable(nc_list[i], dname = dname, fun = mean)
}
max<-do.call(cbind,var_list)
```

End up with two dfs for each variable. Want average per year or per column
```{r}
Tmean <- list()
for (i in 1:ncol(max)) {
  Tmean[[i]] <- ((min[,i] + max[,i]) / 2)
}
Tmean <- do.call(cbind, Tmean)

Tmean <- as.data.frame(Tmean)
colnames(Tmean) <- seq(2007, 2017, 1)
Tmean %<>%
  transmute(TenYrAverage = rowMeans(.)) %>%
  cbind(lonlat, .) %>%
  as.data.table()

# Convert to celcius from Kelvin
Tmean$TenYrAverage <- Tmean$TenYrAverage - 273.15
```

```{r}
saveRDS(Tmean, "../../ncFiles/Processed Data/Tmean")
```

Climate models run in climate_mod_adair.rmd
