---
title: "Reproject Climate Raster"
author: "Max Blasdel"
date: "January 31, 2019"
output: html_document
---

# Purpose: Take output of climate mod script and reproject and resample to match UW data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries

```{r}
library(gdalUtils)
library(raster)
```

# Load original UW dataset, this tif is not used but contains the correct resolution and projection 

```{r}
UW_FCID <- raster("../../data/FCID/UW_FCID_no_wild.tif")

UW_crs <- as.character(crs(UW_FCID))
UW_ext <- bbox(extent(UW_FCID))
tr <- c(30, 30)
```

Reproject the climate raster using nearest neighbor (should preserve values).

A version three is being run with the slightly altered adair equation using 273.15 instead of 273. See climate_mod_adair.rmd for more information on this change.
```{r}
gdalwarp("../../climate_mod/adair_V3.tif",
  "../../climate_mod/climate_V3.tif",
  tr = tr,
  r = "near",
  t_srs = UW_crs,
  te = c(UW_ext),
  overwrite=TRUE)
```

# Reproject other climate models

Find all Q10 tifs
```{r}
q10 <- dir("Sensitivity", pattern = ".tif", full.names = T)
seq <- seq(1.4,2,.2)

# loop through all rasters
for (i in 1:length(q10)) {
  gdalwarp(q10[i],
  paste0("Sensitivity/climate", seq[i], ".tif"),
  tr = tr,
  r = "near",
  t_srs = UW_crs,
  te = c(UW_ext),
  overwrite=TRUE)
}
```

Load and save rasters. This reduces the size of the files considerably
Note - I think the dots in the sequence was causing some problems with the file format
```{r}
cli <- dir("Sensitivity", pattern = "climate", full.names = T)
cli
climate <- raster(cli[1])
climate2 <- raster(cli[2])
climate3 <- raster(cli[3])
climate4 <- raster(cli[4])

# needs to be run indiv for some reason
writeRaster(climate, filename = "Sensitivity/cli_1_4_min.tif")
writeRaster(climate2, filename = "Sensitivity/cli_1_6_min.tif")
writeRaster(climate3, filename = "Sensitivity/cli_1_8_min.tif")
writeRaster(climate4, filename = "Sensitivity/cli_2_min.tif")

```

