---
title: "Decay Implementation"
date: "November 5, 2018"
output:
  pdf_document: default
  html_document: default
  word_document: default
---
This is an augmented version of a script from Micah implementing decay.
Needs to address issue of movement of foliage to duff and incorporation of piled values.

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "../")
```

# Purpose

This document calculates decay over time by size class for harvest and treatment residues. 

# Setup

Load the necessary packages and set ggplot2 theme.

```{r message=FALSE, warning=FALSE}
library(data.table)
library(parallel)
library(ggplot2)
library(tidyverse)

theme_set(theme_classic() + 
                  theme(panel.grid.major = element_line(color = "grey90",
                                                        size = 0.2),
                        strip.background = element_blank()))
```

# Residue 

First, load the treatment raster attribute data, processed in UW_treatments.Rmd.

```{r}
Residue_by_treat <- fread("data/Residue_by_treat.csv")
```

The residue data set is really big because it contains the resdiue values for each FCID and treatment, including the no action treatment. To make developement easier, select a treatment and filter the residue to just that treatment. Additionally, select the first 1,000 FCID, and filter out the rest. **NOTE: This will need to be undone for any final runs.**

```{r}
Residue_by_treat <- Residue_by_treat[Treatment == "RM100"]

Residue_by_treat <- Residue_by_treat[c(1:100),]
```

Load the weighted K values, one for each FCID number. Filter the k values to only include FCIDs that are in the subset of the residue data. Grab only the columns of interest and rename them. *NOTE: Updated with piled materials* **NOTE: updated with non-weighted decay values**

```{r}
decay <- fread("../ResidualTreeLists/decay_FCID_with_climate.csv")

decay <- decay[FCID %in% Residue_by_treat$FCID2018]

decay[is.na(decay)]<-0

decay<-decay[, .(FCID2018 = FCID, # I drop species here and I don't think I need it.
            Species = Species,
            Proportion = proportion,
            Foliage = foliage,
            Fine = FWD,
            Coarse = CWD,
            Pile_Foliage = piled_foliage,
            Pile_Fine = piled_FWD,
            Pile_Coarse = piled_CWD,
            climate = climate)]
```


```{r}
# weighted_decay <- fread("csv/weighted_FCID.csv") 
# 
# weighted_decay <- weighted_decay[FCID2018 %in% Residue_by_treat$FCID2018]
# 
# weighted_decay <- weighted_decay[, .(FCID2018 = FCID2018, 
#                                      Foliage = Weighted_K_foliage,
#                                      Fine = Weighted_K_FWD,
#                                      Coarse = Weighted_K_CWD,
#                                      Pile_Foliage = piled_foliage,
#                                      Pile_Fine = piled_FWD,
#                                      Pile_Coarse = piled_CWD)]
```

Join the k values to the residue data.
```{r}
# This creates multiple entries for each FCID that has multiple species. Amounts will need to be scaled to reflect proportions by species
Residue_by_treat <- merge(Residue_by_treat, decay, by = "FCID2018")
```

Create a simple negative exponential decay function.

```{r}
decay_fun <- function(residue, k_val, t) {
        return(residue * exp(-k_val * t))
}
```

Split the table into a list based on FCID.**12/10**
No issues with multiple entries per FCID. Calculate new amounts of residues based on species abundence of FCID.
```{r}
residue_list <- split(Residue_by_treat, Residue_by_treat$FCID2018)

residue_list <- lapply(residue_list, function(x) x  %>% # Calculates new abundance based on proportions
    mutate(Pulp_6t9_tonsAcre = Pulp_6t9_tonsAcre*Proportion) %>% 
    mutate(Break_6t9_tonsAcre = Break_6t9_tonsAcre*Proportion) %>% 
    mutate(Pulp_4t6_tonsAcre = Pulp_4t6_tonsAcre*Proportion) %>% 
    mutate(Break_4t6_tonsAcre = Break_4t6_tonsAcre*Proportion) %>% 
    mutate(Break_ge9_tonsAcre = Break_ge9_tonsAcre*Proportion) %>% 
    mutate(Branch_tonsAcre = Branch_tonsAcre*Proportion) %>% 
    mutate(Foliage_tonsAcre = Foliage_tonsAcre*Proportion) %>% 
    mutate(Break_4t9_tonsAcre = Break_4t9_tonsAcre*Proportion) %>%
    mutate(Pulp_4t9_tonsAcre = Pulp_4t9_tonsAcre*Proportion))

# residue_list <- residue_list %>% replace_na(0)
```

Max Version
Probably not the fastest, but it works and calculates foliage movement to duff at 50% mass.
Need to run with for non-piled and for piled. Wait until I have climate section done as well. 

```{r}
doDecay <- function(list) {
  test_out <- list()
  for (i in 1:101) { # years 1 through 100
  test_out[[i]] <- unique(list %>% 
    transmute(FCID2018 = FCID2018,
            Treatment = Treatment,
            Year = i-1,
            
            Foliage_time_T = if(sum(decay_fun(Foliage_tonsAcre, Foliage, i-1)) < sum(list$Foliage_tonsAcre*0.5)){ # check for 50% duff
             0# this only sets foliage to zero, need to move to duff
            } else{
              sum(decay_fun(Foliage_tonsAcre, Foliage, i-1))
            }, # Applies exp. decay and sums the results
           
            Branch_time_T = sum(decay_fun(Branch_tonsAcre, Fine, i-1)),
            Pulp4t9_time_T = sum(decay_fun(Pulp_4t9_tonsAcre, Coarse, i-1)),
            Break_4t9_time_T = sum(decay_fun(Break_4t6_tonsAcre, Coarse, i-1)),
            Break_ge9_time_T = sum(decay_fun(Break_ge9_tonsAcre, Coarse, i-1))
            # I want to calculate the duff from the decay loss each year.
            ))
  }
 
  test_out<-do.call(rbind, test_out)
  test_out$annual_duff = 0 # set the initial duff amount to 0
  # removing the pulp4t9 and break4t9 columns since these were doubled up on the residues
  test_out$annual_duff[2:101] = abs(rowSums(test_out[ ,c('Foliage_time_T', 'Branch_time_T', 'Pulp4t9_time_T', 'Break_ge9_time_T')]) %>% diff()*0.02) # calculate annual inputs of duff from decaying material

for (j in 1:100) { # loop to add one time amount of duff from foliage
  if (test_out$Branch_time_T %>% sum() == 0){ # test for 0 rows which cause problems
    break
  if (test_out$Foliage_time_T[j] == 0){
    x<-test_out$Foliage_time_T[j-1]
    test_out$annual_duff[j]<-test_out$annual_duff[j]+x # add the duff, only happens once
    break
   }
  }   
}
  test_out$Duff_time_T<-cumsum(decay_fun(test_out$annual_duff, .002, test_out$Year)) # decay the duff and calculate cumulative totals
    return(test_out)
}

# looping the above function for all residue list items. 
allDecay<-list()
system.time(
for (i in 1:length(residue_list)) {
 allDecay[[i]]<-doDecay(residue_list[[i]])  
}
)
```

Bind into one file
```{r}
allDecay <- do.call(rbind, allDecay)


```



Micah function. I find it confusing and it broke for some reason.
```{r}
start<-Sys.time()
decay_list <- lapply(residue_list, 
                       #mc.cores = detectCores() - 1,
                       function(x){
        out_l <- lapply(seq(1, 100), function(i) {
                # decay each size class
         # orig_mass = x$Foliage_tonsAcre
               decay_i <- x[, .(FCID2018 = FCID2018,
                                Treatment = Treatment,
                                Year = i,
                                Foliage_tonsAcre = decay_fun(Foliage_tonsAcre,
                                                             Foliage, 
                                                             i),
                                Branch_tonsAcre = decay_fun(Branch_tonsAcre,
                                                            Fine,
                                                            i),
                                Pulp_4t9_tonsAcre = decay_fun(Pulp_4t9_tonsAcre,
                                                              Coarse,
                                                              i),
                                Break_4t9_tonsAcre = decay_fun(Break_4t6_tonsAcre,
                                                               Coarse,
                                                               i),
                                Break_ge9_tonsAcre = decay_fun(Break_ge9_tonsAcre,
                                                               Coarse,
                                                               i))]
               # calculate duff addition from each year
               decay_i$Annual_Duff_tonsAcre <- rowSums(decay_i[, c("Foliage_tonsAcre",
                                                                   "Branch_tonsAcre",
                                                                   "Pulp_4t9_tonsAcre",
                                                                   "Break_4t9_tonsAcre",
                                                                   "Break_ge9_tonsAcre")]) * 0.02
                                               
                                               return(decay_i)
                
        })
        
        out_df <- do.call("rbind", out_l)
        
        # decaying the duff at 0.002 percent a year
        # Does not need to be altered
        out_df$Duff_tonsAcre <- decay_fun(cumsum(out_df$Annual_Duff_tonsAcre),
                                          0.002,
                                          out_df$Year)
        
        # save the output after removing old version
                # file.remove(paste0("data/UW/decayed/decayed_residue_",
                #               x,
                #               ".csv")
                # fwrite(decay_i,
                #        paste0("data/UW/decayed/decayed_residue_",
                #               x,
                #               ".csv"))
        
        return(out_df)
})
end<-Sys.time()
end-start
```
Just for reference the above function took about 13.5 hours to run.

Merge the decay data into a single data table.

```{r}
decay <- do.call("rbind", decay_list)
```

Plot some of the output. I randomly selected a FCID.

```{r fig.height=5, fig.width=6}
ggplot(decay[FCID2018 == 102951281], aes(Year, Duff_tonsAcre)) +
        geom_line()

ggplot(decay[FCID2018 == 102951281], aes(Year, Annual_Duff_tonsAcre)) +
        geom_line()

ggplot(decay[FCID2018 == 102951281], aes(Year, Branch_tonsAcre)) +
        geom_line()

ggplot(decay[FCID2018 == 102951281], aes(Year, Foliage_tonsAcre)) +
        geom_line()
```


