---
title: "Mean Decay Values"
author: "Max Blasdel"
date: "September 28, 2018"
output: html_document
---
Goes from the Reorganized species database to the mean species k csv.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document will be used to call the species database, perform some data wrangling, calculate k values for foliage based on angiosperma/gymnosperm distinctions, create distinctions between fine and coarse wood, and add Other hard and softwood values.

Load the desired packages.
```{r include=FALSE}
require(foreign)
require(dplyr)
require(tidyverse)
```

Read in the original species database and inspect.
```{r}
species.db<-read.csv("../data/Reorganized_species_database.csv", stringsAsFactors = F)%>% 
  select(c(K.constant, Genus, Species, Size.Class..cm., Notes, Classification, Fuel.Class))%>%
  rename(size.class = Size.Class..cm.)
head(species.db)
```

I want to calculate affect of species regardless of foliage, but foliage is also a fuel size class. I also want to look at differences between angiosperm and gymnosperm foliage.
```{r}
foliage.gym<- species.db %>%
  filter(Fuel.Class == "foliage", Classification == "Gymnosperm") %>%
  mutate(k.foliage = mean(K.constant)) %>%
  select(Classification, k.foliage) %>%
  distinct()
  
foliage.ang<- species.db %>%
  filter(Fuel.Class == "foliage", Classification == "Angiosperm") %>%
  mutate(k.foliage = mean(K.constant)) %>%
  select(Classification, k.foliage) %>%
  distinct()

foliage<-rbind(foliage.ang, foliage.gym)

head(foliage)
```


I want to add a new column based on the classification for foliage type. 
```{r}
species.db<-inner_join(species.db, foliage, by="Classification")
```


I might want to remove the foliage and snag data before going forward. These values will throw off the overall genus means that will be calculated. 
```{r}
species.db<-species.db %>%
  filter(Fuel.Class != "snag") %>%
  filter(Fuel.Class != "foliage") %>% 
  filter(Fuel.Class != "bark")
```


Make the distinction between CWD and FWD by seperating the data sets based on Fuel Class.
```{r}
FWD<-species.db %>%
  filter(Fuel.Class=="twigs" | Fuel.Class=="branch" | Fuel.Class=="large branch") %>%
#  mutate(class.k = mean(K.constant)) %>% #This is really just for how CWD and FWD as a whole differ
  mutate(Class = "FWD") %>%
  select(-Notes, -Fuel.Class, -size.class)
```


Same step for CWD
```{r}
CWD<-species.db %>%
  filter(Fuel.Class == "log" | Fuel.Class == "large log") %>%
#  mutate(class.k = mean(K.constant)) %>% # Same same
  mutate(Class = "CWD") %>%
  select(-Notes, -Fuel.Class, -size.class)
```


Calculate mean k values for genus species and FWD,CWD distinction
```{r}
fine_coarse<-rbind(FWD,CWD) %>%
  group_by(Genus, Species, Class) %>%
  summarise(mean.k=mean(K.constant)) %>%
  spread(key = Class, mean.k, convert = T)
 # This gives me the genus mean
genus_mean<-species.db %>%
  group_by(Genus) %>%
  transmute(Genus.mean = mean(K.constant)) %>%
  distinct()
# This gives the mean based on species and size distinction with the foliage value and Ang/Gym distinction
fine_coarse<-species.db %>%
  select(Genus, Classification, k.foliage) %>%
  right_join(fine_coarse, by="Genus") %>%
  distinct()
```


Joining the datasets together
```{r}
species.db.update<-left_join(genus_mean,fine_coarse, by="Genus")
```


I want to create values for other hardwood and other softwood. I will use the angiosperm and gymnosperm distinctions here. 
```{r}
ang.gym.mean.values<-species.db %>%
  group_by(Classification) %>%
  summarise(Genus.mean = mean(K.constant)) %>%
  select(-one_of("Classification")) %>%
  mutate(Genus=c("Other hardwood", "Other Softwood")) %>%
  mutate(Species= "Spp.") %>%
  mutate(CWD = NA) %>%
  mutate(FWD = NA)

ang.gym.mean.values<-cbind(ang.gym.mean.values, foliage)

ang.gym.mean.values
```


Linking the dataframes without the species specific information
```{r}
mean.values.all<-rbind.data.frame(species.db.update, ang.gym.mean.values)
```


Writing out the mean decay values by species for later use.
```{r}
write.csv(mean.values.all, file = "../csv/mean.species.k.csv", row.names = F)
```







