---
title: "Deriving missing values for CWD and FWD"
date: "February 25, 2019"
---
# Recreate script to derive missing values known CWD and FWD

# Perform regression analysis to relate decay of CWD to FWD for all species
```{r message=FALSE}
library(tidyverse)
library(magrittr)
```

linear regression equation
```{r}
lm_eqn <- function(df){
    m <- lm(y ~ x, df)
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
         list(a = format(coef(m)[1], digits = 3),
              b = format(coef(m)[2], digits = 3),
             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq))
}
```

```{r}
lm_eqn_noR <- function(df){
    m <- lm(y ~ x, df)
    eq <- substitute(italic(y) == a + b %.% italic(x),
         list(a = format(coef(m)[1], digits = 3),
              b = format(coef(m)[2], digits = 3)))
#             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq))
}
```

```{r}
dat <- read.csv("../csv/mean.species.k.csv", stringsAsFactors = F)
```

Plot out points
```{r}
plot(dat$CWD, dat$FWD)
```

Create df for equation to be displayed
```{r}
df <- dat %>% 
  rename(x = CWD,
         y = FWD)
```


Perform regression equation
```{r}
lm <- lm(FWD ~ CWD, data = dat)
summary(lm)
```

Plotting points with GGplot and with climate normalized values
```{r}
dat %>% 
  ggplot() +
  theme_minimal() +
  geom_point(aes( x = CWD, y = FWD)) +
  geom_abline(slope = lm$coefficients[2], intercept = lm$coefficients[1]) +
  scale_x_continuous(limits = c(0, .2)) +
  geom_text(x = .15, y = .35, label = lm_eqn_noR(df), parse = TRUE) +
  labs(y = "FWD K Constant", x = "CWD K Constant")
```
Output graph
```{r}
ggsave("../images/regression.png", dpi = 300)
```




Applying derived values equation
$$y=0.0393 + 1.2535x$$
Correlation tests
```{r}
cor.test(dat$CWD, dat$FWD, method = "pearson")
```


To derive FWD values 
```{r}
dat %<>% 
  mutate(der.FWD = 0.0393 + CWD*1.2535,
         der.CWD = (FWD-0.0393)/1.2535)

dat
```

Replace occurences of NA with derived value
```{r}
dat %<>% 
  mutate(CWD = coalesce(CWD, der.CWD),
         FWD = coalesce(FWD, der.FWD))

```

Write out for later use.
```{r}
write.csv(dat, "../csv/mean.species.k.Wderived.csv")
```




# Perform regression on recalculated values without climate attached

```{r}
library(readxl)
```

```{r}
dat <- read_xlsx("../../climate modifier/data/Reorganized_species_database_with_climate.xlsx")
```

```{r}
# splits data frame based on size class
groups <- dat %>% 
  mutate(class = case_when(fuel_class == "twigs" | fuel_class == "branch" | fuel_class == "large branch" ~ "FWD",
                           fuel_class == "log" | fuel_class == "large log" ~ "CWD",
                           TRUE ~ "Other")) %>% 
  select(class, k_const, genus, species) %>% 
  mutate(species = paste(genus, species, sep = "_")) %>% 
  select(-genus) %>% 
  group_split(class)
  
dats <- dat %>% 
  mutate(class = case_when(fuel_class == "twigs" | fuel_class == "branch" | fuel_class == "large branch" ~ "FWD",
                           fuel_class == "log" | fuel_class == "large log" ~ "CWD",
                           TRUE ~ "Other"),
         species = paste(genus, species, sep = "_")) %>% 
  select(class, k_const, species) %>% 
  group_by(species, class) %>% 
  mutate(k_const_mean = mean(k_const)) %>% 
  select(-k_const) %>% 
  distinct() %>% 
  spread(key = class, value = k_const_mean, fill = NA) # coerce into long format
```

Prepare data for regression
```{r}
# for lm equations
df <- dats %>% 
  rename(x = CWD,
         y = FWD)

lm <- lm(FWD ~ CWD, data = dats)
```

plot similar to above
```{r}
dats %>% 
  ggplot() +
  theme_minimal() +
  geom_point(aes( x = CWD, y = FWD)) +
  geom_abline(slope = lm$coefficients[2], intercept = lm$coefficients[1]) +
  scale_x_continuous(limits = c(0, .2)) +
  geom_text(x = .15, y = .35, label = lm_eqn_noR(df), parse = TRUE) +
  labs(y = "FWD K Constant", x = "CWD K Constant")
```
Similar results to non-augmented data regression

```{r}
summary(lm)
```

