---
title: "pile_scatter_update"
author: "Max Blasdel"
date: "3/13/2019"
output: html_document
---

# use tidyverse to find percent ground contact and percent above ground

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
library(tidyverse)
library(magrittr)
```

read in data
drop unneeded columns

```{r}
dat <- read.csv("boston.csv", stringsAsFactors = F) %>% 
  select(-c(site,
            type, 
            area_unit, 
            scattered, 
            total_biomass,
            percent_piles))

dat %>% head(n = 7)
```

I think I want to use the gross pile volume as that would have the correct height

```{r}
dat %<>%
  mutate(vol_per_pile = gross_pv/pile_count,
         area_per_pile_sqm = pile_area / pile_count * 10000, # convert ha to meters
         pile_width = sqrt(area_per_pile_sqm / pi) * 2)
```

The equation $$V=(pi*H*W^2)/8$$ describes the volume of the paraboloid. 

Solving for H as $$H=(V*8)/(pi*W^2)$$

```{r}
dat %<>% 
  mutate(pile_height = (vol_per_pile * 8) / (pi * pile_width^2))
dat
```

Calculate amount in ground contact and amount above ground
Based on .5 meter distinction between GC and AG material
Computes to roughly 80/20 split proposed by Van Lear

```{r}
dist_ground = 0.3048

dat %<>% 
  mutate(gc = vol_per_pile - (pi * (pile_height - dist_ground) * pile_width^2)/8,
         ag = vol_per_pile - gc,
         per_gc = gc / (gc + ag) * 100,
         per_ag = ag / (gc + ag) * 100)
```

Summary stats

```{r}
sum_stats <- dat %>% 
  summarise_all(.funs = mean)
sum_stats$per_ag
sum_stats$per_gc
```

Thinking through how to normalize the pile so that one constant can be applied to piled material.

Read in the ag/gc coefficients. Take mean and multiply by percent in above ground class. Add percent in ground contact class times coefficient of 1, since there is no penalty to ground material.

```{r}
agGc <- read.csv("ag_gc_coefficient.csv")

coEf <- agGc %>% 
  filter(ref. == 'Mean') %>% 
  select(ag.gc) %>% 
  unlist()

sum_stats
```

Function for weighted average of piled coefficients
```{r}
piled_k_const <- function(k_const, coEf = 0.721, per_ag = .892, per_gc = .108) {
  
  k_pile <- ((k_const * coEf) * per_ag) + (k_const * per_gc)
  
  return(k_pile)
}

piled_k_const(.09)
```



```{r}
# This is not correct. I want to do a weighted average approach to the piled coefficient
sum_stats %>% 
  transmute(wt_ag = per_ag * coEf,
            wt_gc = per_gc * 1,
            wt_coef = wt_ag + wt_gc)
```

