---
Title: "Decay Figure"
Date: "7/2/2019"
---

# Purpose: Create Decay Figure for Display
Decay figure will show variaition on decay rates through climate for the ten most prominent species in CA

Load libraries
```{r message=FALSE}
options(scipen = 10)
library(raster)
library(tidyverse)
```

Species proportions csv used to determine which species to use for graph

```{r}
proportions <- read.csv("species_props.csv", stringsAsFactors = F)

spec <- read_csv("../ResidualTreeLists/csv/Species.csv", col_types = cols()) %>%
  dplyr::select(c(PlantsCode, FVSCode)) %>%
  rename(Spp = FVSCode)

proportions <- left_join(proportions, spec)
```

# load rasters of decay values for identified species

```{r}
# get names of species
names <- proportions[1:10,] %>% 
  dplyr::select(PlantsCode) %>% 
  t() %>% # transpose
  c() # concatenate

# get all raster locations
rasters <- dir("E:/rasterSpecies/temp_rast_cwd_V3", pattern = ".tif$", full.names = T)

# subset to rasters of interest
rasters <- lapply(rasters, function(x) {
  str_subset(x, pattern = names)
})

# remove empty paths
rasters <- rasters[lapply(rasters, length) > 0]
```

# Load rasters and convert to data.frames
This step may take a long time
```{r}
# define as.data.frame to remove NA values
r.data.frame <- function(x) {
  r <- as.data.frame(x, na.rm = T) %>% 
    mutate(species = names(.)) # add species name identifier
  return(r)
}

rasters <- lapply(rasters, raster)

# call each raster and convert to a data.frame
# may take a really long time
r_list <- lapply(rasters, r.data.frame)

rbind_list(r_list) %>% 
  saveRDS("rasterValues.RDS")
```

# inspect output
Output may be too large. 
Data is in wrong format and needs to be cleaned up by removing NA values
Read in data, clean, and write out as indivual dataframes to join later on

```{r}
names <- names %>% str_subset("PSME|ABCO",negate = T)

for (i in 1:length(names)) {
  data <- readRDS("rasterValues.RDS") %>% 
    filter(species == names[i]) %>% 
    select(species, everything())
  
  data <- data[,colSums(is.na(data)) == 0] %>% 
    rename(value = 2)
    
  paste(data[1,1])
  
  saveRDS(data, paste0("decay_by_species/", data[1,1], ".RDS"))
}
```


# Load final data
Get all data frames and join into one long df

```{r}
dfs <- dir("decay_by_species", full.names = T)

dfs <- lapply(dfs, read_rds)

dfs <- bind_rows(dfs)

#saveRDS(dfs, "rasterValuesAll.RDS")
```

# plotting

Subset data for plot construction

```{r}
dfs <- read_rds("rasterValuesAll.rds")

# take random sample for quicker iterations in constructin plots 
# dfs <- dfs[sample(nrow(dfs), 3000, replace = F), ]

# clear gc() to free memory
gc()
gc()
```

Prepare data

```{r}
# arrange by species
dfs <- dfs %>% arrange(species)

# add class distinction
dfs <- dfs %>% 
  mutate(class = case_when(species == "QUKE" ~ "Angiosperm",
                           species == "QUCH2" ~ "Angiosperm",
                           TRUE ~ "Gymnosperm"))


# add names to df to make more readable
dfs <- left_join(dfs, proportions %>% select(PlantsCode, name), by = c("species" = "PlantsCode"))

# drop species column for space
dfs <- dfs %>% 
  select(-species)

# reorder by factor to match % prominence
dfs$name <- factor(dfs$name, levels = c("Douglas-fir",
                                        "Ponderosa Pine", 
                                        "White Fir", 
                                        "Incense Cedar", 
                                        "Jeffrey Pine",
                                        "California Black Oak",
                                        "California Red Fir",
                                        "Sugar Pine",
                                        "Canyon Live Oak",
                                        "Grand Fir"), ordered = T)
dfs %>% head()
gc()
```

*consider* Taking a sample of the data as the whole dataset is too large and not a lot may be gained from looking at everything as opposed to a smaple

Plotting

```{r}
ggplot(dfs) +
  geom_boxplot(aes(name, value, fill = class)) +
  theme_minimal() +
  labs(y = "Decay Values",
       x = "Species",
       fill = "Class") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


consider using a groub_by_all %>% count to determine hwo many rows I could possibly remove by using some type of factor. I have lots of duplicate rows, but I want to retain the same ratio of unique values. 

```{r}
memory.limit()

t <- dfs %>% group_by_all() %>% count()

gc()
t %>% arrange(name)


t$n %>% sum()

ggplot(t) +
  geom_boxplot(aes(name, value))
```

