---
title: "Species Proportions from Residual Tree List"
Date: "March 22, 2019"
---

# Purpose

Create graphics showing species proportions for California based on total DBH

```{r message=FALSE}
library(tidyverse)
library(data.table)
```

Using residual tree list from remove NA scenario as base data

```{r}
dat <- fread("../ResidualTreeLists/csv/rmna.csv") %>% 
   select(-c(OBJECTID, TrtId, LCR, MCW))

spec<-fread("../ResidualTreeLists/csv/Species.csv") %>%
  select(c(PlantsCode, FVSCode, CommonName, ScientificName, Classification)) %>%
  rename(Spp = FVSCode)
spec
```

Read in the residual tree list and parse it down to proportional DBH grouped by FCID.

```{r}
dat <- dat %>%
  group_by(FCID2018, Spp) %>%
  #group_by(Spp) %>%
  summarise(sumDBH = sum(DBH)) %>%
  ungroup() %>%
  group_by(FCID2018) %>%
  mutate(proportion = sumDBH/sum(sumDBH))
```

Join more species data

```{r}
dat <- left_join(dat, spec, by = "Spp")
```

Calculate proportions values

```{r}
options(scipen = 20)
dat %>% head()

dat %>% 
  group_by(Spp) %>% 
  transmute(sumDBH = sum(sumDBH),
            name = CommonName,
            class = Classification) %>% 
  distinct() %>% 
  ungroup() %>% 
  mutate(prop = sumDBH/sum(sumDBH) * 100) %>% 
  arrange(desc(prop))# %>% 
#  write.csv("species_props.csv") # wrote out for later use
```

#Processed Data

Read in data
```{r}
dat <- read.csv("species_props.csv", stringsAsFactors = F, row.names = 'X')
dat %>% head()
```

Simple function for changing names to capitalized
```{r}
simpleCap <- function(x) {
  s <- strsplit(x, " ")[[1]]
  paste(toupper(substring(s, 1,1)), substring(s, 2),
      sep="", collapse=" ")
}

dat$name<- sapply(dat$name, simpleCap)
```

Thinking about some plots for the data

```{r}
dat$name <- dat$name %>% factor(levels = dat$name)

dat %>% 
  filter(prop >=1) %>% 
  ggplot() +
  theme_classic() +
    geom_bar(aes(x = name, y= prop, fill = class), stat = "identity") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  ggtitle("Proportion of Total DBH of California Species") +
  scale_fill_discrete(name= "Classification") +
  xlab(NULL) + # include title here or not?
  ylab("Proportion (%)")
```

```{r}
ggsave("species_proportions.png", dpi = 300)
```

